---
title: "Santa Barbara Rainfall"
author: "Lindsay Hafen"
date: "11/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(lubridate)
library(chron)
```

## Read in data 
- Use here(): data is in data file 
- Change to lower_snake_case to make coding easier using janitor's clean_names()
- Data was downloaded as standard instead of metric, so the measurements will be in 100th's of inches

```{r}
sb_rainfall_raw <- read_csv(here("data", "santa_barbara_rainfall.csv")) %>% 
  clean_names() 
```

## Look at data summary statistics

```{r}
summary(sb_rainfall_raw)
```



## Wrangle data
- Organize the date
- Create a subset with one year of data for second plot

```{r}
sb_rainfall <- sb_rainfall_raw %>% 
  separate(col = date, 
           into = c("date", "time"),
           sep = " ") %>% 
  mutate(year = year(date),
         month = month(date)) %>% 
  filter(year < 2014, year > 1971)

sb_pcp_year <- sb_rainfall %>% 
  group_by(year)

fillrow = which(sb_pcp_year$qpcp < 0 | sb_pcp_year$qpcp > 800)
fillrow
sb_pcp_year$qpcp[fillrow] = NA
summary(sb_pcp_year)

sb_pcp_sum <- sb_pcp_year %>% 
  group_by(year) %>% 
  summarise(total_pcp = sum(qpcp, na.rm = TRUE))
```

## Make a plot of the whole range
- Plot qpcp against year

```{r}
ggplot(sb_pcp_sum, aes(y = total_pcp, x = year)) +
  geom_col() +
  labs(x = "Year", y = "Total precipitation 1/100 (in)", title = "Santa Barbara Precipitation (1972-2013)")
```

## Make 5 graphs of 1 year periods

```{r}
fillrow = which(sb_rainfall$qpcp < 0 | sb_rainfall$qpcp > 800)
fillrow
sb_rainfall$qpcp[fillrow] = NA
```

```{r}
sb_pcp_short_1 <- sb_rainfall %>% 
  filter(year == 2009) %>% 
  filter(qpcp > 0) %>% 
  mutate(date_new = ymd(date)) %>% 
  group_by(month) %>% 
  summarise(tot_pcp = sum(qpcp, na.rm = TRUE))

ggplot(sb_pcp_short_1, aes(x = month, y = tot_pcp)) +
  geom_col() +
  scale_x_continuous(breaks = seq(1, 12, 1)) +
  labs(x = "Month", y = "Precipitation in 1/100 (in)", title = "Santa Barbara Rainfall (2009)")
```

```{r}
sb_pcp_short_2 <- sb_rainfall %>% 
  filter(year == 2010) %>% 
  #filter(qpcp > 0) %>% 
  mutate(date_new = ymd(date)) %>% 
  group_by(month) %>% 
  summarise(tot_pcp = sum(qpcp, na.rm = TRUE))

ggplot(sb_pcp_short_2, aes(x = month, y = tot_pcp)) +
  geom_col() +
  scale_x_continuous(breaks = seq(1, 12, 1)) +
  labs(x = "Month", y = "Precipitation in 1/100 (in)", title = "Santa Barbara Rainfall (2010)")
```

```{r}
sb_pcp_short_3 <- sb_rainfall %>% 
  filter(year == 2011) %>% 
  #filter(qpcp > 0) %>% 
  mutate(date_new = ymd(date)) %>% 
  group_by(month) %>% 
  summarise(tot_pcp = sum(qpcp, na.rm = TRUE))

ggplot(sb_pcp_short_3, aes(x = month, y = tot_pcp)) +
  geom_col() +
  scale_x_continuous(breaks = seq(1, 12, 1)) +
  labs(x = "Month", y = "Precipitation in 1/100 (in)", title = "Santa Barbara Rainfall (2011)")
```

```{r}
sb_pcp_short_4 <- sb_rainfall %>% 
  filter(year == 2012) %>% 
  #filter(qpcp > 0) %>% 
  mutate(date_new = ymd(date)) %>% 
  group_by(month) %>% 
  summarise(tot_pcp = sum(qpcp, na.rm = TRUE))

ggplot(sb_pcp_short_4, aes(x = month, y = tot_pcp)) +
  geom_col() +
  scale_x_continuous(breaks = seq(1, 12, 1)) +
  labs(x = "Month", y = "Precipitation in 1/100 (in)", title = "Santa Barbara Rainfall (2012)")
```

```{r}
sb_pcp_short_5 <- sb_rainfall %>% 
  filter(year == 2013) %>% 
  #filter(qpcp > 0) %>% 
  mutate(date_new = ymd(date)) %>% 
  group_by(month) %>% 
  summarise(tot_pcp = sum(qpcp, na.rm = TRUE))

ggplot(sb_pcp_short_5, aes(x = month, y = tot_pcp)) +
  geom_col() +
  scale_x_continuous(breaks = seq(1, 12, 1)) +
  labs(x = "Month", y = "Precipitation in 1/100 (in)", title = "Santa Barbara Rainfall (2013)")
```

## Calculate monthly averages

```{r}
fillrow = which(sb_rainfall$qpcp < 0 | sb_rainfall$qpcp > 800)
fillrow
sb_rainfall$qpcp[fillrow] = NA
```


```{r}
sb_totals <- sb_rainfall %>% 
  group_by(year, month) %>% 
  summarize(total_pcp = sum(qpcp, na.rm = TRUE))

sb_averages <- sb_totals %>% 
  group_by(month) %>% 
  summarize(average_pcp = mean(total_pcp))
```

```{r}
ggplot(sb_averages, aes(x = month, y = average_pcp)) +
  geom_col() +
  labs(x = "Month", y = "Precipitation 1/100 (in)", title = "Average Santa Barbara Rainfall (1972-2013)")
```





